Introduction to\\ Swedish Civil Contingencies Agency (MSB) methodological support for introducing \ac{isms}.

### 2018-04-09 msbframework/msbintro/msbintro.tex
added gdpr to legal requirements, divided one slide to two slides for readability
Small fixes, spelling errors etc.

\begin{frame}{MSB}
  \begin{itemize}
    \item Swedish Civil Contingencies Agency (MSB).
    \item \enquote{MSB is responsible for issues concerning civil protection,
        public safety, emergency management and civil defence, as long as no
        other authority has responsibility\cite[About MSB]{msbse}\@.
        .}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Informationssäkerhet.se}
  \begin{itemize}
    \item MSB ran a project called SVISA\@: 'Stöd för Verksamheters 
      InformationsSäkerhetsArbete'.

@@

\begin{frame}{eu:gdpr}{General Data Protection Regulation}
  \begin{itemize}
    \item Replacing Personal Data Act 1998:204
    \item Regulates how personal data are allowed to be stored and used.
      \begin{itemize}
        \item Protects a persons right of privacy\cite[art. 5]{eu:gdpr}
        \item Personal data can only be stored for as long as there is a need
          for it\cite[art. 5]{eu:gdpr}.
      \item Each person must give explicit
        consent to a company/organisation to store, publish or share personal
        data about them\cite[art. 7]{eu:gdpr}.
        \item Ensure that a person have the right to get access to all personal
          data that has been collected about them\cite[art. 15]{eu:gdpr}.
        \item Persons right to erasure\cite[art. 17]{eu:gdpr}. 
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
    \begin{itemize}
      \item Possibility to transfer personal data (data
          portabililty)\cite[art. 20]{eu:gdpr}.  
      \item Records regarding processing activities must be maintained \cite[art.
        30]{eu:gdpr}.
      \item Personal data breaches must be reported to the supervisory
        authority, without undue delay, not later than 72 hours\cite[art.
        33]{eu:gdpr}.
    \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
\begin{itemize}
  \item Data protection by design and default \cite[art. 25]{eu:gdpr}.
      \item Individuals needs to be informed about the personal data
        breach\cite[art. 34]{eu:gdpr}.
    \item An independent agency should ensure that these rules are followed
      \cite[art 51]{eu:gdpr}.
    \item If the organisation does not comply to these regulations, they can be
      fined up to 4\% worldwide turnover or €20 Million, whichever is greater
      \cite[art 83]{eu:gdpr}.
    \item and much more.
\end{itemize}
\end{frame}
\begin{frame}{MSBFS 2016:1}
  \begin{itemize}
    \item Due to increased electronic information exchange in the society, there
      is now demands put on how governmental agency work with information
      security.

@@

  Detta innebär bland annat att en myndighet måste: 
  \begin{enumerate}
    \item upprätta en informationssäkerhetspolicy och andra styrande dokument 
      som behövs för myndighetens informationssäkerhet,
    \item utse en eller flera personer som leder och samordnar arbetet med 
      informationssäkerhet,
    \item klassificera sin information med utgångspunkt i krav på 
      konfidentialitet, riktighet och tillgänglighet,
    \item utifrån risk- och sårbarhetsanalyser och inträffade incidenter 
      avgöra hur risker ska hanteras, samt besl uta om åtgärder för 
      myndighetens informationssäkerhet,
    \item dokumentera granskningar och säkerhetsåtgärder av större betydelse 
      som har vidtagits.
    \item 
  \end{enumerate}
\end{frame}

@@

\begin{frame}{Legal requirements}
  Agreement, laws and regulations.
  \begin{itemize}
    \item Personal Data Act,
    \item The Archives Act,
    \item Public Access to Information and Secrecy Act
    \item MSBFS 2016:1,
    \item Security Act.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Risk matrix}{Consequences}
  \begin{itemize}
    \item How severe is the consequence for the organisation of the threat
      happens?
    \item Clarify who will be a affected: Organisation through a side effect of
      consequences in the society? 
    \item Makes it easier to give examples for the different consequence levels.
      For example: loss of reputation, higher cost, \dots.
  \end{itemize}
  \begin{center}
    Severe -- Considerable -- Moderate -- Negligible
  \end{center}
\end{frame}

@@

\begin{frame}{Risk management}
  \begin{itemize}
    \item Decide whether the identified threats should be rectified or be
      accepted:
      \begin{itemize}
        \item Accept,
        \item eliminate,
        \item transfer,
        \item rectify.
      \end{itemize}
    \item What measures should be taken?
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{MSB}
  \begin{itemize}
    \item Swedish Civil Contingencies Agency (MSB).
    \item \enquote{MSB is responsible for issues concerning civil protection,
        public safety, emergency management and civil defence, as long as no
        other authority has responsibility\cite[About MSB]{msbse}\@.
        .}
    \item This includes protection of our information systems.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Informationssäkerhet.se}
  \begin{itemize}
    \item In 2011 MSB ran a project called SVISA\@: 'Stöd för Verksamheters 
      InformationsSäkerhetsArbete'.

@@

\begin{frame}{GDPR}{General Data Protection Regulation}
  \begin{itemize}
    \item Replacing Personal Data Act 1998:204
    \item Regulates how personal data are allowed to be stored and used.
      \begin{itemize}
        \item Protects a persons right of privacy\cite[art. 5]{eu:gdpr}
        \item Personal data can only be stored for as long as there is a need
          for it\cite[art. 5]{eu:gdpr}.
        \item Each person must give explicit consent to a company/organisation
          to store, publish or share personal data about them\cite[art.
          7]{eu:gdpr}.  
        \item Ensure that a person have the right to get access
          to all personal
          data that has been collected about them\cite[art. 15]{eu:gdpr}.
        \item Persons right to erasure\cite[art. 17]{eu:gdpr}. 
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
    \begin{itemize}
      \item Possibility to transfer personal data (data
          portabililty)\cite[art. 20]{eu:gdpr}.  
      \item Data protection by design and default \cite[art. 25]{eu:gdpr}.
      \item Records regarding processing activities must be maintained \cite[art.
        30]{eu:gdpr}.
      \item Personal data breaches must be reported to the supervisory
        authority, without undue delay, not later than 72 hours\cite[art.
        33]{eu:gdpr}.
    \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
\begin{itemize}
    \item Individuals needs to be informed about the personal data
        breach\cite[art. 34]{eu:gdpr}.
    \item An independent agency should ensure that these rules are followed
      \cite[art 51]{eu:gdpr}.
    \item If the organisation does not comply to these regulations, they can be
      fined up to 4\% worldwide turnover or €20 Million, whichever is greater
      \cite[art 83]{eu:gdpr}.
    \item and much more.
\end{itemize}
\end{frame}
\begin{frame}{MSBFS 2016:1}
  \begin{itemize}
    \item Due to increased electronic information exchange in the society, there
      is now demands put on how governmental agency work with information
      security.

@@

\end{frame}
\begin{frame}{MSBFS:2016:1}
  Detta innebär bland annat att en myndighet måste: 
  \begin{enumerate}
    \item upprätta en informationssäkerhetspolicy och andra styrande dokument 
      som behövs för myndighetens informationssäkerhet,
    \item utse en eller flera personer som leder och samordnar arbetet med 
      informationssäkerhet,
    \item klassificera sin information med utgångspunkt i krav på 
      konfidentialitet, riktighet och tillgänglighet,
    \item utifrån risk- och sårbarhetsanalyser och inträffade incidenter 
      avgöra hur risker ska hanteras, samt besl uta om åtgärder för 
      myndighetens informationssäkerhet,
    \item dokumentera granskningar och säkerhetsåtgärder av större betydelse 
      som har vidtagits.
  \end{enumerate}
\end{frame}

@@

\begin{frame}{Legal requirements}
  Agreement, laws and regulations.
  \begin{itemize}
    \item Personal Data Act,
    \item General Data Protection Regulation
    \item The Archives Act,
    \item Public Access to Information and Secrecy Act
    \item MSBFS 2016:1,
    \item Security Act.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Risk matrix}{Consequences}
  \begin{itemize}
    \item How severe is the consequence for the organisation if the threat
      happens?
    \item Clarify who will be a affected
    \item Makes it easier to give examples for the different consequence levels.
      For example: loss of reputation, higher cost, \dots.
  \end{itemize}
  \begin{center}
    Severe -- Considerable -- Moderate -- Negligible
  \end{center}
\end{frame}

@@

\begin{frame}{Risk management}
  \begin{itemize}
    \item Decide whether the identified threats should be rectified or be
      accepted:
      \begin{itemize}
        \item Accept,
        \item Eliminate,
        \item Transfer,
        \item Rectify.
      \end{itemize}
    \item What measures should be taken?
  \end{itemize}
\end{frame}
